---
title: "Exploratory Data Analysis for Water Table Depth"
author: "Madhup Kumar"
date: "December 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse, quietly = T)
library(plotly, quietly = T)
```


Exploratory data analysis of quarterly data for water table depth in Haryana for the period 2012-2016.

## Loading Data

Loading data

```{r depthload}
depth_db <- read.csv("water_table_depth_quarterly_2015_2016.csv")
depth_db$Quarter <- ordered(depth_db$Quarter, levels=c("May 2015", "August 2015","November 2015","January 2016"))
depth_db$Depth <- as.double(as.character(depth_db$Depth))

summary(depth_db)
```


```{r}

# Replace 0 to NA for Depth
depth_db$Depth[depth_db$Depth==0] <- NA
depth_db$Depth[depth_db$Depth==99.000] <- NA
str(depth_db)
```


The dataset has water table depth (in meters) averaged over all wells of a Block. The depth is measured in the months of May, August, November and January.

##Missing Data

```{r}
missingDepth <- prop.table(table(is.na(depth_db$Depth)))

```
`r missingDepth[2]`% missing values.

IMPORTANT: How to tackle missing data. Best approach here will be to check against well-wise data.

## Counts

Number of records per qtr and per district
```{r}
table(depth_db$Quarter)
table(depth_db$District)
```

```{r}
prop.table(table(depth_db$District))
```




###Outlier Analysis

Box plot of all qtrs
```{r}
ggplotly(ggplot(data=depth_db[,], aes(x=Quarter, y=Depth, fill=Quarter)) + geom_boxplot())
```

###Average Depth by Qtr

```{r}
avg_depth_qtr <- depth_db %>% group_by_(.dots=c("Quarter")) %>% dplyr::summarize(Mean = mean(Depth, na.rm = T))
avg_depth_qtr <- as.data.frame(avg_depth_qtr)
avg_depth_qtr
```

For all qtrs
```{r fig.width=10,fig.height=50}
ggplotly(ggplot(data = depth_db[,], aes(x=District, y=Depth, fill=District)) + geom_boxplot() + ylim(0,50) + facet_grid(Quarter~.) + ylab("Quarterly Average Depth"))

```


IMPORTANT: Outliers need to be tackled using well-wise depth data (which is also available)

