---
title: "Exploratory Data Analysis for Agriculture Yield"
author: "Abhishek Sinha"
date: "December 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr, quietly = T)
library(ggplot2, quietly = T)
library(plotly, quietly = T)
```


Exploratory data analysis of crop yield in Haryana for the period 2012-2016.

## Loading Data

Loading yield data

```{r yieldload}
yield_db <- read.csv("Tidy_Data_final.csv")
summary(yield_db)
```


```{r}
yield_db$Yield <- as.integer(as.character(yield_db$Yield))
yield_db$Year <- as.factor(yield_db$Year)

# Replace 0 to NA for Yield
yield_db$Yield[yield_db$Yield==0] <- NA

str(yield_db)
```


The dataset has yield (produced/output in tonnes per hectare). The data is granual upto Block level and can be traced by Year and Crop Type.

##Missing Data

```{r}
missingYield <- prop.table(table(is.na(yield_db$Yield==0))) * 100

```
`r missingYield[1]`% missing values in yield date.

IMPORTANT: How to tackle missing data. Best approach here will be remove these rows.

## Crop wise Yield

Number of records per crop
```{r}
table(yield_db$Crop)
```

```{r}
prop.table(table(yield_db$Crop)) *100
```

The data is quite balanced.


###Outlier Analysis

Box plot of all crops excluding Sugarcane
```{r}
ggplotly(ggplot(data=yield_db[yield_db$Crop!='Sugarcane',], aes(x=Crop, y=Yield, fill=Crop)) + geom_boxplot())
```

Box plot Sugarcane
```{r}
ggplotly(ggplot(data=yield_db[yield_db$Crop=='Sugarcane',], aes(x=Crop, y=Yield, fill=Crop)) + geom_boxplot())
```

Looking at boxplot infomration, it looks yield of Sugarcane is on a totally different scale compared to other Crops.

*Talk to the team about this.*


###Average Yield by Crop

```{r}
avg_yield_crop <- yield_db %>% group_by_(.dots=c("Crop","Year")) %>% dplyr::summarize(Mean = mean(Yield, na.rm = T))
avg_yield_crop <- as.data.frame(avg_yield_crop)
avg_yield_crop
```

For all crops excluding Sugarcane
```{r fig.width=10,fig.height=11}
ggplotly(ggplot(data = yield_db[yield_db$Crop!='Sugarcane',], aes(x=Year, y=Yield, fill=Year)) + geom_boxplot() + facet_grid(Crop~., scales = "free_y") + ylab("Average Yield"))
```

For Sugarcane

```{r}
ggplotly(ggplot(data = yield_db[yield_db$Crop=='Sugarcane',], aes(x=Year, y=Yield, fill=Year)) + geom_boxplot() + ylab("Average Yield"))
```


IMPORTANT: Outliers need to be tacked by crop and year

