---
title: "Soil Nutrient EDA"
author: "Abhishek Sinha"
date: "January 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ggplot2)
library(plotly)
```

## Loading Data

```{r cache=TRUE}
soil_df <- fread("../Soil_Nutrient_Year.csv", header = T, stringsAsFactors = F, showProgress = T)
soil_df$V1 <- NULL
```

## Summary of Soil Nutrient Data Set

```{r}
summary(soil_df)
```
## PH


```{r}
table(is.na(soil_df$SoilPh))
table(soil_df$SoilPh>14) # 1652 records have incorrect data (<1%)
```

6319 is missing data (<1%).  
1652 rows have incorrect data, >14.

Average value: `r mean(soil_df$SoilPh, na.rm = T)`

```{r}
ggplot(data=soil_df[soil_df$SoilPh<=14,], aes(x=Year, y=SoilPh, fill=as.factor(Year))) + geom_boxplot()
```


## Electrical Conductivty

```{r}
table(is.na(soil_df$ElectricalConductivity)) 
```

7917 missing value in EC (~1%). 

```{r}
ggplot(data=soil_df, aes(x=as.factor(Year), y=ElectricalConductivity, fill=as.factor(Year))) + geom_boxplot()
```


Looking at the historgram

```{r}
ggplot(soil_df) + aes(x = ElectricalConductivity) + geom_histogram(bins = 5, colour ="black", fill="#FF6666")
```


## Organic Compund

Missing values
```{r}
table(is.na(soil_df$OrganicCarbon))
```

1093 missing vales. Mean value: `r mean(soil_df$OrganicCarbon, na.rm=T) `.

OC above 1 vs less than 1
```{r}
table(soil_df$OrganicCarbon<=1)
```

Only `r 100*(4850/(679535+4850))`% of OC value is above 1. Some treatment is required.

```{r}
temp <- soil_df[soil_df$OrganicCarbon<=1,]
ggplot(data=temp, aes(x=as.factor(Year), y=OrganicCarbon, fill=as.factor(Year))) + geom_boxplot()
rm(temp)
```


## Nitrogen

Missing Values
```{r}
table(is.na(soil_df$Nitrogen))
```

Majority of data is missing for Nitrogen. (`r 659896*100/(659896+25582)`%)

HYPOTHESIS: Organic Compound (OC) can be used as the proxy variable for Nitrogen.

### Bi-variate analysis of OC and N
```{r}
data_nit_oc <- soil_df[is.na(soil_df$Nitrogen)==F & is.na(soil_df$OrganicCarbon)==F,]
summary(lm(Nitrogen ~ OrganicCarbon, data_nit_oc))
```


Correlation value: `r cor(data_nit_oc$Nitrogen, data_nit_oc$OrganicCarbon)`.  

Both R Square and Correlation clearly suggests that Nitrogen Value and OC values are not correlated.

Clearly, the hypothesis is incorrect.

With a very high missing values for Nitrogen, we can easily remove the variable.


## Phosphorous

Missing Values
```{r}
table(is.na(soil_df$Phosphorous))
```

Low Missing Value `r 100*2639/(685478)`%.

Mean `r mean(soil_df$Phosphorous, na.rm = T) `.

Box Plot
```{r}
ggplot(data=soil_df[is.na(soil_df$Phosphorous)==F,], aes(x=as.factor(Year), y=Phosphorous, fill=as.factor(Year))) + geom_boxplot()
```

```{r}
upper_lmt <- 14.49 + (1.5*IQR(soil_df$Phosphorous, na.rm = T)) #Taken from Summary
table(soil_df$Phosphorous<=upper_lmt)
```

Approximately, 11% of above the outlier range, which is very high. 

Box Plot on log transformation
```{r}
ggplot(data=soil_df[is.na(soil_df$Phosphorous)==F,], aes(x=as.factor(Year), y=log(Phosphorous), fill=as.factor(Year))) + geom_boxplot()
```

The data looks normally distributed on log transformation. 


## Potassium

Missing Values
```{r}
table(is.na(soil_df$Potassium))
```

Low Missing Value `r 100*2643/(685478)`%.

Mean `r mean(soil_df$Potassium, na.rm = T) `.

Box Plot
```{r}
ggplot(data=soil_df[is.na(soil_df$Potassium)==F,], aes(x=as.factor(Year), y=Potassium, fill=as.factor(Year))) + geom_boxplot()
```

Box Plot on log transformation
```{r}
ggplot(data=soil_df[is.na(soil_df$Potassium)==F,], aes(x=as.factor(Year), y=log(Potassium), fill=as.factor(Year))) + geom_boxplot()
```

Log transformation is highly normal distributed.

## Sulphur

Missing Values
```{r}
table(is.na(soil_df$Sulphur))
```

Low Missing Value `r 100*30037/(685478)`%.

Mean `r mean(soil_df$Sulphur, na.rm = T) `.

Box Plot
```{r}
ggplot(data=soil_df[is.na(soil_df$Sulphur)==F,], aes(x=as.factor(Year), y=Sulphur, fill=as.factor(Year))) + geom_boxplot()
```

Box Plot on log transformation
```{r}
ggplot(data=soil_df[is.na(soil_df$Sulphur)==F,], aes(x=as.factor(Year), y=log(Sulphur), fill=as.factor(Year))) + geom_boxplot()
```

Log transformation is normal distributed with a lttle skewness on left side.

## Zinc

Missing Values
```{r}
table(is.na(soil_df$Zinc))
```

Low Missing Value `r 100*5764/(685478)`%.

Mean `r mean(soil_df$Zinc, na.rm = T) `.

Box Plot
```{r}
ggplot(data=soil_df[is.na(soil_df$Zinc)==F,], aes(x=as.factor(Year), y=Zinc, fill=as.factor(Year))) + geom_boxplot()
```

Lot of outliers.

Box Plot on log transformation
```{r}
ggplot(data=soil_df[is.na(soil_df$Zinc)==F,], aes(x=as.factor(Year), y=log(Zinc), fill=as.factor(Year))) + geom_boxplot()
```

Log transformation is normal distributed.


## Iron


Missing Values
```{r}
table(is.na(soil_df$Iron))
```

Low Missing Value `r 100*5647/(685478)`%.

Mean `r mean(soil_df$Iron, na.rm = T) `.

Box Plot
```{r}
ggplot(data=soil_df[is.na(soil_df$Iron)==F,], aes(x=as.factor(Year), y=Iron, fill=as.factor(Year))) + geom_boxplot()
```

Again lot of outliers.

Box Plot on log transformation
```{r}
ggplot(data=soil_df[is.na(soil_df$Iron)==F,], aes(x=as.factor(Year), y=log(Iron), fill=as.factor(Year))) + geom_boxplot()
```

Log transformation is normal distributed.


## Copper

Missing Values
```{r}
table(is.na(soil_df$Copper))
```

Very high Missing Value `r 100*352636/(685478)`%. 

NEED TO DISCUSS


Mean `r mean(soil_df$Copper, na.rm = T) `.

Box Plot
```{r}
ggplot(data=soil_df[is.na(soil_df$Copper)==F,], aes(x=as.factor(Year), y=Copper, fill=as.factor(Year))) + geom_boxplot()
```

Again lot of outliers.

Box Plot on log transformation
```{r}
ggplot(data=soil_df[is.na(soil_df$Copper)==F,], aes(x=as.factor(Year), y=log(Copper), fill=as.factor(Year))) + geom_boxplot()
```

Log transformation is normal distributed.

## Magnesium

Missing Values
```{r}
table(is.na(soil_df$Magnesium))
```

Missing Value of `r 100*9963/(685478)`% is a little high but impuation should be fine. NEED TO DISCUSS.

Mean `r mean(soil_df$Magnesium, na.rm = T) `.

Box Plot
```{r}
ggplot(data=soil_df[is.na(soil_df$Magnesium)==F,], aes(x=as.factor(Year), y=Magnesium, fill=as.factor(Year))) + geom_boxplot()
```

Again lot of outliers.

Box Plot on log transformation
```{r}
ggplot(data=soil_df[is.na(soil_df$Magnesium)==F,], aes(x=as.factor(Year), y=log(Magnesium), fill=as.factor(Year))) + geom_boxplot()
```

Log transformation is normal distributed.

## Boron

Missing Values
```{r}
table(is.na(soil_df$Boron))
```

Almost all values are missing. Missing Value in `r 100*680505/(685478)`% of data. NEED TO REMOVE THIS feature.